---
title: "Quiz 1"
output: pdf_document
date: '2022-09-02'
---

```{r, include = FALSE}
library(ggplot2)
library(dplyr)
```



1. Create a data frame with x and y columns. X values have to be 6, and Y's values must be
numbers from 1 to 10. In the final part, create a Z column by multiplying X values to the Y
to get the multiplication table of 6. (1 point)

```{r}
df <- data.frame(x=6, y=1:10)
df["z"] <- df["x"] * df["y"]
df

```

2. Load the data mpg.

```{r}

data(mpg)

```

2.1 Check the structure of the mpg data. How many character and numeric columns are there? (0.5 point)

```{r}
str(mpg)
```

There are 6 character columns and 5 numeric columns (including integer columns)


2.2 How many cars have class of minivan with year of production 2008. (1 point)

```{r}
mpg %>% filter(class=="minivan", year==2008) %>% count()
```


2.3 Please select car which has: (1 point)

- year of production 2008
- model a4 quattro
- displ 3.1 
- cty 17 
return value of column trans

```{r}
(mpg %>% filter(year == 2008, model == "a4 quattro", displ == 3.1, cty == 17))$trans
```



2.4 Calculate the average hwy for each manufacturer.What is the average hwy for car manufacturer jeep. (1 point)

```{r}
grped <- mpg %>% group_by(manufacturer) %>% summarize(mean=mean(hwy))
grped %>% filter(manufacturer == "jeep")

```


2.5 Return first 5 rows of the previous question. (0.5 point)

```{r}
head(grped)
```


2.6 Please select calculate average cty for each manufactuer then return manufacturers audi,volkswagen,toyota and jeep (1 point)


```{r}
grped <- mpg %>% group_by(manufacturer) %>% summarize(mean=mean(cty))
grped %>% filter(manufacturer %in% c("audi", "volkswagen", "toyota", "jeep"))
```



2.7 Please select rows from 10 to 30 and columns 1,4,5 and 6 (1 point)

```{r}
mpg[10:30, c(1,4,5,6)]
```


2.8. Calculate difference between hwy and cty and plot the histogram of the new feature. (1 point)
```{r}
hist(mpg$hwy - mpg$cty)
```


Overall: 7 points


