---
title: "Quiz 2"
output:
  pdf_document: default
  html_document: default
date: '2022-09-12'
---
---
title: "Quiz2"
output: html_document
date: '2022-09-11'
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)

```

```{r}
data(iris)
```


1.1. Create a scatterplot of Sepal width and Sepal length. (0.5 point)

```{r}
ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length)) + geom_point()

```

1.2. Add regression line using function geom_smooth() and specify parameters where method = "lm" and se = FALSE. What relationship can you see? (0.5 point)

```{r}
ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length)) + geom_point() + geom_smooth(method="lm", se=FALSE)
```
negative correlation and slope

2.1.Create a scatterplot of Sepal width and Sepal length. Specify the 3rd color as Species.(1 point)

```{r}
ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length, color=Species)) + geom_point()
```

2.2. Change the color of legend specifying colors red, violet and blue. (1 point)

```{r}
ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length, color=Species)) + geom_point() + scale_color_manual(values=c("red", "violet", "blue"))
```

2.3. Add Regression line to your plot using geom_smooth() and specify parameters where method = "lm" and se = FALSE. What relationship you can see now? (1 point)

```{r}
ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length, color=Species)) + geom_point() + scale_color_manual(values=c("red", "violet", "blue")) + geom_smooth(method="lm", se=FALSE)

```
Positive correlation and slope for each species

3. Use faceting for the species category. And specify scales parameter to free. (1 point)

```{r}

ggplot(data = iris, aes(x=Sepal.Width, y=Sepal.Length, color=Species)) + geom_point() + scale_color_manual(values=c("red", "violet", "blue")) + geom_smooth(method="lm", se=FALSE) + facet_wrap(Species ~ .)
```






